A spectral-Galerkin turbulent channel flow solver for large-scale simulations

### 2017-01-13 paper.tex
Tikz figure
Submitted

%\newcommand{\mmo}[1]{\emph{#1}}

@@

\begin{abstract}
A fully (pseudo-)spectral solver for direct numerical simulations of large-scale turbulent channel flows is described. The solver utilizes the Chebyshev base functions suggested by J. Shen [SIAM J. Sci. Comput., 16, 1, 1995], that lead to stable and robust numerical schemes, even at very large scale. New algorithms for direct solution of the linear systems are devised, and we show that the solver is accurate and applicable for very high Reynolds numbers. Specifically, the solver is shown to reproduce the first order statistics of Hoyas and Jim\'{e}nez [Phys. Fluids, 18(1), 2006], for a channel flow at $Re_{\tau}=2000$, which is higher than has been reported by a fully spectral method before. The solver is available through the open source project spectralDNS [https://github.com/spectralDNS].
\end{abstract}
\begin{keyword}
DNS \sep Fourier \sep Chebyshev \sep Biharmonic \sep Helmholtz \sep turbulence
\end{keyword}

@@

The largest known channel simulations to date have been performed by Lee and Moser \cite{leemoser15}, where, for $Re_{\tau}\approx 5200$, they used a computational box of resolution $[10240 \times 1530 \times 7680]$ for the streamwise, wall-normal and spanwise directions, respectively. Lee and Moser used seventh-order B-splines for the wall-normal direction. Other simulations of similar magnitude have been performed by Hoyas and Jim\'{e}nez \cite{hoyas06, hoyas08}, Lozano-Duran and Jim\'{e}nez \cite{Lozano2014} and Bernardini, Pirozzoli and Orlandi \cite{bernardini2014}. Bernardini et al. used second-order finite differences throughout. The Jim\'{e}nez group used dealiased Fourier in the two periodic directions and seventh-order compact finite differences, with fourth-order consistency and extended spectral-like resolution \cite{Lele92}, for the wall-normal direction. The solver by Jim\'{e}nez' group is reported to switch from Chebyshev to finite differences if the resolution is above a certain threshold \cite{hoyas08} (reached around $Re_{\tau}=1000$). In other words, they attempt to use fully spectral methods for as large $Re_{\tau}$ as possible. As previously mentioned, spectral methods are attractive for their accuracy and resolution properties, that are superior to those of any finite difference or spline method. As such, it is desirable to develop fully spectral solvers that can be used for large-scale turbulence simulations.

In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to construct Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to very fast direct solvers. To the authors knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised. Shen's bases have been used for the Navier Stokes equations before, though. Bouchon et.al.  \cite{Bouchon01} describe a spectral-Galerkin formulation very similar to the one used in this paper. However, they choose the Legendre basis over Chebyshev, neglecting the higher cost of transforms and arguing instead for the simpler structure of the matrices and the lack of weight functions for the Legendre basis. The cost of the Legendre transforms, though, makes the basis unsuitable for larger scales, and the simulations performed in \cite{Bouchon01} were rather small.

In this paper we will describe and assess a spectral-Galerkin channel flow solver based on Fourier and Shen's Chebyshev basis \cite{Shen95}. We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS} and we will describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise, in Sec~\ref{sec:implementation}. We will finally show in, Sec~\ref{sec:verification}, that roundoff is not a major issue, and that the Shen-Fourier spectral-Galerkin method is in deed applicable to large-scale simulations, larger than what has been reported for fully spectral channel simulations before. The solver has been implemented in the open source code spectralDNS \cite{spectralDNS}, where the bulk of the solver is written in high-level Python \cite{python}, with critical parts of the code migrated to Cython \cite{cython} for efficiency.

\section{Preliminaries}
\label{sec:prelim}
The Navier-Stokes equations, used to describe turbulent flow in a doubly 
periodic channel, can be written in rotational form as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{P}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time, and the nonlinearity $\bm{\mathcal{H}}(\bm{x}, t) = (\mathcal{H}_x, \mathcal{H}_y, \mathcal{H}_z) = \bm{u}\times 
\bm{\omega}$, where $\bm{\omega} = \nabla \times \bm{u}$.  The constant dynamic viscosity is denoted as $\nu$ 
and $P(\bm{x}, t)$ is a pressure modified to account for both the driving force, $\beta(t)$, and the kinetic energy, i.e., $P = p + \bm{u} \cdot \bm{u}/2 + \beta y$, where $p$ is the instantaneous 
pressure normalized by a constant density. The computational domain is 
$\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$, with channel walls 
located at $x=\pm 1$, such that no-slip applies at 
$ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions with periodic lengths $L_y$ and $L_z$, respectively.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall-normal direction. Three different sets of basis functions and function spaces are relevant for the wall-normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & V_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x}, \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{V}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2}, \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
\frac{k+1}{k+3} T_{k+4}(x), & \N{V}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4}, \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
suggested by Shen \cite{Shen95}, and satisfy, respectively, the boundary conditions 
$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. 
%The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 1) = 0 \}$.

Three-dimensional basis functions and function spaces, that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$ and 
\begin{align}
K_N = \Big\{&\bm{k} \in \mathcal{Z}\times\mathcal{R}^2 / (l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
 &(l, m, n) \in  [0,1, \ldots, N_x] \times  [-\frac{N_y}{2},-\frac{N_y}{2}+1,\ldots, 
 \frac{N_y}{2}-1] \times [-\frac{N_z}{2},-\frac{N_z}{2},+1,\ldots, \frac{N_z}{2}-1] \Big\} 
 \label{eq:wavenumbermesh}
\end{align}
is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
notation $K^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
dimensions of $K_N$. Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers, for the different bases, in the wall-normal direction.

The computational mesh in real space can be created using $N = (N_x, N_y, 
N_z)$ intervals, where the two periodic directions use uniform intervals. The 
computational mesh is given as
\begin{align}
  X_N = \Big\{ \bm{x} \in \mathcal{R}^3/&(x_i, y_j, z_k) \in \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
  &(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\}, \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents
\begin{equation}
 h(i) = \begin{cases}
   \cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Chebyshev-Gauss-Lobatto points}, \\
   \cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Chebyshev-Gauss points}. \\
 \end{cases}
\end{equation}

For the Navier Stokes equations we look for solutions of the velocity 
components of the form
\begin{align}
u(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{u}_{\bm{k}}(t) 
\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solx} \\
v(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{v}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_soly} \\
w(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{w}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solz}
\end{align}
where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion 
coefficients for the velocity component in $x$-direction (and similar for the 
other two components) and the scaling by $N_y$ and $N_z$ is merely for 
convenience 
and compliance with the definition used for the inverse discrete Fourier 
transform. Note that from now on we will simply use the notation $\hat{u}$ for 
$\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. 
Likewise we will simply use $u$ for $u(\bm{x}, t)$.

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast inverse transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n,t)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath 
\underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))) \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, and $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse discrete
Fourier transforms along directions $y$ and $z$ respectively. For simplicity, we introduce here a special notation called the inverse 
\emph{Shen} transform, $\N{\mathcal{S}}^{-1}_x$, that is used to transform coefficients from spectral to physical space in a series expansion that is using either one of the bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}). The inverse Shen transform, $\mathcal{S}^{-1}_x$, is performed along the wall-normal 
$x$ direction, and it may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}), see Alg~\ref{alg:ifst} in the Appendix. The transforms are slightly different for the three 
bases and 
${\mathcal{S}}_x^{-1}, \D{\mathcal{S}}_x^{-1}$ and $\N{\mathcal{S}}_x^{-1}$ are used 
to distinguish between them with obvious notation.

A three-dimensional scalar product in the weighted $L^2_w(\Omega)$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, 
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for 
periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
this work we will make use of the discrete weighted $l^2_w(\Omega)$ space, 
where quadrature is employed for the integration. As such, the scalar product 
corresponds to sums that may be computed using fast transforms. Using 
$\N{\psi}_{\bm{k}}$ for $v$ we redefine the scalar product with quadrature as
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))) \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three-dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is used to represent the forward \emph{Shen} scalar product in the 
$x$-direction. The weights, $w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Chebyshev-Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Chebyshev-Gauss points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$, see Sec. 1.11 of \cite{kopriva09}.%\footnote{Note that $c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. (2.7) of \cite{Shen95}), but for the discrete $l^2_w(\Omega)$ it follows that $c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}

>>>>

\usepackage{tikz}
\usepackage{pgfplots}

%\newcommand{\mmo}[1]{\emph{#1}}
\interfootnotelinepenalty=10000
\newcommand*{\Scale}[2][4]{\scalebox{#1}{$#2$}}%
\newcommand{\half}{\Scale[0.6]{\tfrac{1}{2}}}
\newcommand{\nhalf}{\Scale[0.6]{-\tfrac{1}{2}}}

@@

\begin{abstract}
A fully (pseudo-)spectral solver for direct numerical simulations of large-scale turbulent channel flows is described. The solver utilizes the Chebyshev base functions suggested by J. Shen [SIAM J. Sci. Comput., 16, 1, 1995], that lead to stable and robust numerical schemes, even at very large scale. New and fast algorithms for the direct solution of the linear systems are devised, and algorithms and matrices for all required scalar products and transforms are provided. We validate the solver for very high Reynolds numbers. Specifically, the solver is shown to reproduce the first order statistics of Hoyas and Jim\'{e}nez [Phys. Fluids, 18(1), 2006], for a channel flow at $Re_{\tau}=2000$. The solver is available through the open source project spectralDNS [https://github.com/spectralDNS].
\end{abstract}
\begin{keyword}
DNS \sep Fourier \sep Chebyshev \sep Biharmonic \sep Helmholtz \sep turbulence
\end{keyword}

@@

The largest known channel simulations to date have been performed by Lee and Moser \cite{leemoser15}, where, for $Re_{\tau}\approx 5200$, they used a computational box of resolution $[10240 \times 1530 \times 7680]$ for the streamwise, wall-normal and spanwise directions, respectively. Lee and Moser used seventh-order B-splines for the wall-normal direction. Other simulations of similar magnitude have been performed by Hoyas and Jim\'{e}nez \cite{hoyas06, hoyas08}, Lozano-Duran and Jim\'{e}nez \cite{Lozano2014} and Bernardini, Pirozzoli and Orlandi \cite{bernardini2014}. Bernardini et al. used second-order finite differences throughout. The Jim\'{e}nez group used dealiased Fourier in the two periodic directions and seventh-order compact finite differences, with fourth-order consistency and extended spectral-like resolution \cite{Lele92}, for the wall-normal direction. The solver by Jim\'{e}nez' group is reported to switch from Chebyshev to finite differences if the resolution is above a certain threshold \cite{hoyas08} (reached around $Re_{\tau}=1000$). In other words, they attempt to use a fully spectral discretization for as large $Re_{\tau}$ as possible. As previously mentioned, spectral methods are attractive for their accuracy and resolution properties, that are superior to those of any finite difference or spline method. As such, it is desirable to develop fully spectral solvers that can be used for large-scale turbulence simulations.

In his seminal papers \cite{Shen94,Shen95}, Jie Shen describe how to construct Legendre and Chebyshev basis functions that lead to sparse matrices, susceptible to very fast direct solvers. To the author's knowledge, the bases have not been used for large-scale channel flow simulations, and algorithms for the required direct solvers have not, until now, been devised. Shen's bases have been used for the Navier Stokes equations before, though. Bouchon et.al.  \cite{Bouchon01} describe a spectral-Galerkin formulation very similar to the one used in this paper. However, they choose the Legendre basis over Chebyshev, which has some consequences when aiming for large-scale, because fast transforms are required in moving from spectral to physical space, and back again. For Fourier and Chebyshev bases, the Fast Fourier Transforms (FFTs of $\mathcal{O} (N \log N)$) apply directly. However, until recently, the discrete Legendre transforms required $\mathcal{O}(N^2)$ operations. This scaling has now been improved by several authors, as recently reviewed by Hale and Townsend \cite{Hale14, Hale2015}, but the methods are still not quite on par with the FFTs. For example, Hale and Townsend describe an $\mathcal{O}(N (\log N)^2 / \log \log N)$ algorithm, using intermediate fast transforms from Legendre to Chebyshev coefficients.

In this paper we will describe and assess a spectral-Galerkin channel flow solver based on Fourier and Shen's Chebyshev basis \cite{Shen95}. The solver will consist of parts that scale at worst as $N \log N$, for a 1D problem of size $N$, and as such as $N^3 \log N$ for a 3D box of size $N^3$. We will give a proper description of the theoretical basis in Sec~\ref{sec:prelim}, the discretization of Navier-Stokes equations in Sec~\ref{sec:discretizationNS}, and we will describe necessary algorithms, including a new fast direct solver for the biharmonic problem that arise, in Sec~\ref{sec:implementation}. We will finally show, in Sec~\ref{sec:verification}, that roundoff is not a major issue, and that the Shen-Fourier spectral-Galerkin method is in deed applicable to large-scale simulations. The solver has been implemented in the open source code spectralDNS \cite{spectralDNS}, where the bulk of the code is written in high-level Python \cite{python}, with critical parts migrated to Cython \cite{cython} for efficiency.

\section{Basis functions and fast transforms}
\label{sec:prelim}
The Navier-Stokes equations, used to describe turbulent flow in a doubly 
periodic channel, can be written in rotational form as
\begin{align}
 \frac{\partial \bm{u}}{\partial t}   &= \bm{\mathcal{H}} + \nu 
 \nabla^2 \bm{u} - \nabla{\tilde{p}}, \notag \\
 \nabla \cdot \bm{u} &= 0, \label{eq:NS}
\end{align}
where $\bm{u}(\bm{x}, t)=(u, v, w)$ is the velocity vector, $\bm{x}=(x, y, z)$ 
and $t$ are position and time, and the nonlinearity $\bm{\mathcal{H}}(\bm{x}, t) = (\mathcal{H}_x, \mathcal{H}_y, \mathcal{H}_z) = \bm{u}\times 
\bm{\omega}$, where $\bm{\omega} = \nabla \times \bm{u}$.  The constant dynamic viscosity is denoted as $\nu$ 
and $\tilde{p}(\bm{x}, t)$ is a pressure modified to account for both the driving force, $\beta(t)$, and the kinetic energy, i.e., $\tilde{p} = p + \bm{u} \cdot \bm{u}/2 + \beta y$, where $p$ is the instantaneous 
pressure normalized by a constant density. The computational domain is 
$\Omega=[-1, 1]\times [0, L_y] \times [0, L_z]$, with channel walls 
located at $x=\pm 1$, such that no-slip applies at 
$ \bm{u}(\pm 1, y, z, t) = 0$. The walls are spanning 
the $y-z$ plane and the equations are periodic in the $y$ and $z$ directions with periodic lengths $L_y$ and $L_z$, respectively.

The domain $\Omega$ is discretized using $N = (N_x, N_y, N_z)$ intervals, where the two periodic directions use uniform intervals. The computational mesh is given as
\begin{align}
X_N = \Big\{ \bm{x} \in \mathbb{R}^3 | &(x_i, y_j, z_k) = \left( h(i), \frac{jL_y}{N_y}, \frac{kL_z}{N_z} \right), \text{where} \notag \\
&(i, j, k) \in [0, 1, \ldots, N_x] \times [0, 1, \ldots, N_y-1] \times [0, 1, \ldots, N_z-1] \Big\}, \label{eq:Xn}
\end{align}
where $x_i = h(i)$ represents
\begin{equation}
h(i) = \begin{cases}
\cos \left(\frac{i \pi }{N_x} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Chebyshev-Gauss-Lobatto points}, \\
\cos \left(\frac{(2i +1)\pi}{2N_x+2} \right) \, &\forall \, i=0,1, \ldots, N_x \quad  \text{for Chebyshev-Gauss points}. \\
\end{cases}
\end{equation}

The spectral-Galerkin method makes use of a three-dimensional scalar product in the weighted $L^2_{\sigma}(\Omega)$ space, that is defined as
\begin{align}
\left<u, \upsilon \right>_{\sigma} &= \int_{\Omega} {u(\bm{x}) \upsilon^*(\bm{x})} \sigma(\bm{x})\,dxdydz, 
\end{align}
where $\upsilon^*$ is the complex conjugate of the test function $\upsilon$ and the weights $\sigma$ are unity for 
periodic directions and  $\sigma(x)=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
this work we will make use of the discrete weighted $l^2_{\sigma}(\Omega)$ space, 
where quadrature is employed for the integration. As such, we redefine the scalar product as
\begin{equation}
\left<u, \upsilon\right>_{\sigma} = \sum_{i=0}^{N_x}\sum_{j=0}^{N_y-1} \sum_{k=0}^{N_z-1} u(x_i, y_j, z_k) \upsilon^*(x_i, y_j, z_k) \sigma(x_i), \label{eq:quadrature}
\end{equation}
that is more amendable to the fast transforms that will be defined later in this section.

The Navier Stokes equations are discretized using Fourier basis functions for the periodic directions, and a combination of Chebyshev polynomials in the wall-normal direction. Three different sets of basis functions and function spaces are relevant for the wall-normal direction
\begin{align}
&  \phi_k(x) = T_k(x), & W_{N_x} &= \text{span}\{\phi_k\}_{k=0}^{N_x}, \label{eq:Tk}\\
& \D{\phi}_k(x) = T_k(x) - T_{k+2}(x), & \D{W}_{N_x} &= \text{span} \{ \D{\phi}_k\}_{k=0}^{N_x-2}, \label{eq:phiD}\\
& \N{\phi}_k(x) = T_k(x) - \frac{2(k+2)}{k+3} T_{k+2}(x) + 
\frac{k+1}{k+3} T_{k+4}(x), & \N{W}_{N_x} &= \text{span} \{\N{\phi}_k\}_{k=0}^{N_x-4}, \label{eq:phiN} 
\end{align}
where $T_k(x)$ is the $k$'th degree Chebyshev polynomial of the first kind. The 
basis functions and function spaces in (\ref{eq:phiD}) and (\ref{eq:phiN}) were 
suggested by Shen \cite{Shen95}, and satisfy, respectively, the boundary conditions 
$\D{\phi}_k(\pm 1) = 0$, $\N{\phi}_k(\pm 1)=0$ and $\N{\phi}'_k(\pm 1)=0$. 
%The last two function spaces may alternatively be written as $\D{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1)=0 \}$ and $\N{V}_{N_x} = \{v \in V_{N_x}: v(\pm 1) = v'(\pm 1) = 0 \}$.

Three-dimensional basis functions and function spaces, that are periodic in $y$ 
and $z$ directions, can now be defined as
\begin{align}
  \psi_{\bm{k}}(\bm{x}) = \phi_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad V_N &= \text{span} \{ \psi_{\bm{k}}(\bm{x}):\, \bm{k} \in K_N  \}, \\
  \D{\psi}_{\bm{k}}(\bm{x}) = \D{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \D{V}_N &= \text{span} \{ \D{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \D{K}_N  \}, \\
  \N{\psi}_{\bm{k}}(\bm{x}) = \N{\phi}_{l}(x)e^{ \imath(\underline{m} y + \underline{n} z)}, \quad \N{V}_N &= \text{span} \{ \N{\psi}_{\bm{k}}(\bm{x}):\, \bm{k} \in \N{K}_N  \},
\end{align}
where $\imath=\sqrt{-1}$. 
%\begin{align}
%K_N = \Big\{&\bm{k} \in \mathbb{Z}\times\mathbb{R}^2 / (l, \underline{m}, \underline{n}) \in \left( l, \frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
% &(l, m, n) \in  [0,1, \ldots, N_x] \times  [-\frac{N_y}{2},-\frac{N_y}{2}+1,\ldots, 
% \frac{N_y}{2}-1] \times [-\frac{N_z}{2},-\frac{N_z}{2},+1,\ldots, \frac{N_z}{2}-1] \Big\} 
% \label{eq:wavenumbermesh}
%\end{align}
%is the wavenumber mesh for $V_N$. The other two wavenumber meshes $\D{K}_N$ and 
%$\N{K}_N$ are both equal to $K_N$ except from in index $l$ that 
%ranges $[0, 1, \ldots, N_x-2]$ and $[0, 1, \ldots, N_x-4]$ for $\D{K}_N$ 
%and $\N{K}_N$ respectively, see (\ref{eq:phiD}) and (\ref{eq:phiN}). We use 
%notation $K^p(m, n)=K_N(0, m, n)$ when referring to only the two periodic 
%dimensions of $K_N$. Likewise, ${K}^x(l)={K}_N(l, 0, 0), \D{K}^x(l)=\D{K}_N(l, 0, 0)$ and $\N{K}^x(l)=\N{K}_N(l, 0, 0)$ refer to the wavenumbers, for the different bases, in the wall-normal direction.
The wavenumbermesh, $K_N$, for space $V_{N}$, is defined by the Cartesian product of wavenumbers from the two periodic and the inhomogeneous wall-normal direction: $K_N(l,m,n)= K^x(l) \times K^p(m,n)$, where
\begin{align}
K^p = \Big\{&(\underline{m}, \underline{n}) = \left(\frac{2 \pi m}{L_y}, \frac{2 \pi n}{L_z} \right), \text{where} \notag \\
&(m, n) \in  [-\frac{N_y}{2},-\frac{N_y}{2}+1,\ldots, 
\frac{N_y}{2}-1] \times [-\frac{N_z}{2},-\frac{N_z}{2},+1,\ldots, \frac{N_z}{2}-1] \Big\} 
\label{eq:wavenumbermeshp}
\end{align}
and $K^x(l) = \{l \in \mathbb{Z} | l=0,1, \ldots, N_x\}$. The two remaining wavenumber meshes, $\D{K}_{N}$ and $\N{K}_{N}$, differ from $K_N$ only in the range of the first index sets, $\D{K}^x$ and $\N{K}^x$, ending in $N_x-2$ and $N_x-4$, respectively (see Eqs. (\ref{eq:phiD}) and (\ref{eq:phiN})).

In the spectral-Galerkin method we look for solutions of the velocity 
components of the form
\begin{align}
u(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \N{K}_N} \hat{u}_{\bm{k}}(t) 
\N{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solx} \\
v(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{v}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_soly} \\
w(\bm{x}, t) &= \frac{1}{N_yN_z}\sum_{\bm{k} \in \D{K}_N} \hat{w}_{\bm{k}}(t) 
\D{\psi}_{\bm{k}}(\bm{x}), \label{eq:u_solz}
\end{align}
where $\hat{u}_{\bm{k}}(t) = \hat{u}(l, {m}, {n}, t)$ are the expansion 
coefficients for the velocity component in $x$-direction (and similar for the 
other two components) and the scaling by $N_y$ and $N_z$ is merely for 
convenience 
and compliance with the definition used later for the inverse discrete Fourier 
transform. Note that from now on we will simply use the notation $\hat{u}$ for 
$\hat{u}_{\bm{k}}(t)$, when it is possible to simplify without loss of clarity. 
Likewise we will simply use $u$ for $u(\bm{x}, t)$.

For an efficient method it is crucial to be able to compute $u$ quickly from $\hat{u}$, or, vice versa, to compute $\hat{u}$ quickly from $u$. This may be achieved using the fast transform methods to be defined next.
Consider first how to compute $u$ from the known expansion coefficients $\hat{u}$. Writing out the summation terms, the expression (\ref{eq:u_solx}) may be 
evaluated on the entire mesh (\ref{eq:Xn}) as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n,t)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath 
\underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))),
  \label{eq:ifft} 
\end{align}
where $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform in space $\N{V}_N$, and $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse discrete
Fourier transforms along directions $y$ and $z$ respectively. For simplicity, we have introduce here a special notation called the inverse 
\emph{Shen} transform, ${\mathcal{S}}^{-1}_x$, that is used to transform coefficients from spectral to physical space in a series expansion that is using either one of the bases in (\ref{eq:Tk}, \ref{eq:phiD}, \ref{eq:phiN}). The inverse Shen transform, $\mathcal{S}^{-1}_x$, is performed along the wall-normal 
$x$ direction, and it may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}), see Alg~\ref{alg:ifst} in the Appendix. The transforms are slightly different for the three 
bases and 
${\mathcal{S}}_x^{-1}, \D{\mathcal{S}}_x^{-1}$ and $\N{\mathcal{S}}_x^{-1}$ are used 
to distinguish between them with obvious notation. Similarly, $\D{\mathcal{T}}^{-1}$ and ${\mathcal{T}}^{-1}$ define the inverse transforms for spaces $\D{V}_N$ and $V_N$. Note that a transform in  any one direction is performed over all indices in the other two directions. For example, for the transform in the $x$-direction we have
\begin{equation}
u(x_i,m,n,t) = \N{\mathcal{S}}_x^{-1}(\hat{u}(l,m,n,t)) \quad \forall \, i \in [0,1,\ldots, N_x] \text{ and }  m,n \in {K}^p,
\end{equation}
and similar for the other two directions.
